**Evalution of binary classification pipelines and methods for 16S rRNA gene data**

### 2018-11-19 submission/manuscript.Rmd
add sklearn version
versions

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
```

@@

######################################################################
#---Load .tsv data with cvAUCs for hyper-parameters  ----#
######################################################################
# Read in hyper-parameter AUCs generated from L2 logistic regression model for all samples:
logit <- read.delim('../data/process/L2_Logistic_Regression_parameters.tsv', header=T, sep='\t')%>%
  select(-X) %>% 
  mutate(meanAUC=rowMeans(.[, 2:101])) %>% 
  select(C, meanAUC) 
# Read in hyper-parameter AUCs generated from L1 Linear SVM model for all samples:
l1svm <- read.delim('data/process/L1_SVM_Linear_Kernel_parameters.tsv', header=T, sep='\t')%>%
  select(-X) %>% 
  mutate(meanAUC=rowMeans(.[, 2:101])) %>% 
  select(C, meanAUC) 
# Read in hyper-parameter AUCs generated from L2 Linear SVM model for all samples:
l2svm <- read.delim('data/process/L2_SVM_Linear_Kernel_parameters.tsv', header=T, sep='\t')%>%
  select(-X) %>% 
  mutate(meanAUC=rowMeans(.[, 2:101])) %>% 
  select(C, meanAUC)

@@

@@

Colorectal cancer is one of the leading cause of death among cancers in the United States. Each person in the industrialized world has on average a one-in-twenty chance of developing colorectal cancer (CRC) in their lifetime and once diagnosed, more than one-third will not survive 5 years. Colonoscopy as a screening tool is very effective, however it is very invasive, expensive and have a low rate of patient adherence. Therefore, there is a need for improved non-invasive methods to screen individuals. One proposed non-invasive screening tool is using gut microbiome-based biomarkers. Patients with colorectal cancer have different stool community of microbes compared to adults with normal colons. This difference however cannot be explained by a single or a handful of features in the gut microbiome but by many of them in relation to one another. Therefore, machine learning is a great tool to investigate the differences between the gut microbiomes of CRC patients and healthy individuals. Previous studies have shown that human hemoglobin levels and bacterial population abundances in the stool help us predict screen relevant growth in the colon, however the literature for the problem of classification colorectal cancer diagnosis vary greatly, with areas under the receiver operating characteristic curve (AUC) of 0.7-0.8. The variation in classification performance is based in part on differences in the task definition, in
part on differences in the study populations, and in part on the evaluation method. The highest
reported AUCs were from studies of ... Additionally, some of the reported results were not obtained from testing on held-out sets. In this study, we have defined classification pipelines with L2-regularized logistic regression, L1 and L2 linear suppor vector machines (SVM), radial basis function SVM , decision tree, random forest and XGBoost. We evaluated the generalization and prediction performance of these methods. We also compared each method based on their reproducibility, robustness, actionability, interpretibility and susceptibility to overfitting.

@@

#### Data definitions and pre-processing
We use FIT results and OTU abundances as features. However these are in different scales. We use MinMaxScaler to transforms features by scaling each feature to a [0-1] range.

@@

```{r}
Table1
Table2
Table3
```

>>>>

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
library(reticulate)
use_python("/Library/Frameworks/Python.framework/Versions/3.6/bin/python3", required = TRUE)
```

@@

######################################################################
#---Load .tsv data with cvAUCs for hyper-parameters  ----#
######################################################################
# Read in hyper-parameter AUCs generated from L2 logistic regression model for all samples:
logit <- read.delim('../data/process/L2_Logistic_Regression_parameters.tsv', header=T, sep='\t')%>%
  select(-X) %>% 
  mutate(meanAUC=rowMeans(.[, 2:101])) %>% 
  select(C, meanAUC) 
# Read in hyper-parameter AUCs generated from L1 Linear SVM model for all samples:
l1svm <- read.delim('../data/process/L1_SVM_Linear_Kernel_parameters.tsv', header=T, sep='\t')%>%
  select(-X) %>% 
  mutate(meanAUC=rowMeans(.[, 2:101])) %>% 
  select(C, meanAUC) 
# Read in hyper-parameter AUCs generated from L2 Linear SVM model for all samples:
l2svm <- read.delim('../data/process/L2_SVM_Linear_Kernel_parameters.tsv', header=T, sep='\t')%>%
  select(-X) %>% 
  mutate(meanAUC=rowMeans(.[, 2:101])) %>% 
  select(C, meanAUC)

@@

```{python, cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
import platform
python_version = platform.python_version()
import sklearn
sklearn_module = sklearn.__version__
```

@@

Colorectal cancer is one of the leading cause of death among cancers in the United States. Each person in the industrialized world has on average a one-in-twenty chance of developing colorectal cancer (CRC) in their lifetime and once diagnosed, more than one-third will not survive 5 years. Colonoscopy as a screening tool is very effective, however it is very invasive, expensive and have a low rate of patient adherence. Therefore, there is a need for improved non-invasive methods to screen individuals. One proposed non-invasive screening tool is using gut microbiome-based biomarkers. Patients with colorectal cancer have different stool community of microbes compared to adults with normal colons. This difference however cannot be explained by a single or a handful of features in the gut microbiome but by many of them in relation to one another. Therefore, machine learning is a great tool to investigate the differences between the gut microbiomes of CRC patients and healthy individuals. Previous studies have shown that human hemoglobin levels and bacterial population abundances in the stool help us predict screen relevant growth in the colon, however the literature for the problem of classification colorectal cancer diagnosis vary greatly, with areas under the receiver operating characteristic curve (AUC) of 0.7-0.8. The variation in classification performance is based in part on differences in the task definition, in
part on differences in the study populations, and in part on the evaluation method. The highest
reported AUCs were from studies of ...  In this study, we have defined classification pipelines with L2-regularized logistic regression, L1 and L2 linear suppor vector machines (SVM), radial basis function SVM , decision tree, random forest and XGBoost. We evaluated the generalization and prediction performance of these methods. We also compared each method based on their reproducibility, robustness, actionability, interpretibility and susceptibility to overfitting.

@@

#### Data definitions and pre-processing
We use FIT results and OTU abundances as features. However these are in different scales. We use python v`r paste(py$python_version)`, module scikit-learn v`r paste(py$sklearn_module)` function MinMaxScaler() to transform features by scaling each feature to a [0-1] range.

@@
